---
title: "hair_cortisol_dog_combined"
author: "Alex German"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Check working directory
```{r}
getwd()
```





# Load packages
```{r}
library(dlookr)
library(dplyr)
library(ggplot2)
library(psych)
library(reshape)
library(readxl)
library(tidyverse)
library(vioplot)
library(vtable)

library(brms)
library(bayesplot)
library(bayestestR)
library(loo)
library(priorsense)
library(rethinking)

```


# Load data
```{r}
df <- read_xlsx("hair_cort_dog_all.xlsx", col_types = c("text", "text",  
                               "text", "text", "text", "text",
                               "text", "numeric","text", "skip",
                               "numeric", "skip", "skip", 
                               "numeric", "skip"))
df <- as.data.frame(df)
```



# INITIAL DATA PLOTTING AND EXPLORATION

## Check characteristics of df
```{r}
dim(df) # will tell you how many rows and columns the dataset has
class(df) # will tell you what data structure has the dataset been assigned
```


## Explore the dataset to understand its structure.
```{r}
head(df)
```


# 1.  Get summary stats for numeric data ####
```{r}
numeric_df <- Filter(is.numeric, df)
describe(numeric_df) # the describe function in psych provides summary stats
```



# 2. Check normality of all numeric variables
## a. graphical assessment
```{r}
plot_normality(numeric_df)
```


## b. shapiro-wilk test
```{r}
apply(numeric_df, 2, shapiro.test)
```


## c. repeat Q-Q plots with transformed data
### i. log(cortisol)
```{r}
qqnorm(df$cortisol)
qqline(df$cortisol, col = "red")
qqnorm(log(df$cortisol))
qqline(log(df$cortisol), col = "red")
```

### ii Shapiro test for log cortisol
```{r}
shapiro.test(log(df$cortisol))
```



# 3. Check cortisol data numerically
## a. summarisecortisol and log cortisol
```{r}
summary(df$cortisol)
sd(df$cortisol)
summary(log(df$cortisol))

```


## b. log-transform cortisol variable
```{r}
df$lgCort <- log(df$cortisol)
summary(df$lgCort)
sd(df$lgCort)
```


## c. visualise lgCort variable
```{r}
hist(df$lgCort)
```


## d. Check hair cortisol by comorbidity
### i. cortisol
```{r}
describeBy(cortisol ~ comorbidity, data = df)
```

## ii. lgCort
```{r}
describeBy(lgCort ~ comorbidity, data = df)
```


## iii. Calculate mean difference in lgCort for different comorbidity status
```{r}
diffLgCortCo <- mean(df$lgCort[df$comorbidity == "yes"]) - mean(df$lgCort[df$comorbidity == "no"])
diffLgCortCo
exp(diffLgCortCo)
exp(0.25)
diffLgCortCo / 1.29
```


# 4. Generate variables for use in modelling

## a. create a binary outcome variable for group, where "stopped" is ref category
```{r}
df$weight_loss_success <- df$group
df$weight_loss_success <- ifelse(df$weight_loss_success == "completed", "yes", "no")
```


## b. create simple category name for breed and convert to factor
```{r}
df$breed <- df$breed_group
df$breed <- factor(df$breed, levels = c("mix", "ckcs", "pug", "ret", "other"))
head(df$breed)
```

## c. reorder season so spring is reference and rest are in orderd
```{r}
df$season <- factor(df$season, levels = c("spring", "summer", "autumn", "winter"))
head(df$season)
```


## d. make light hair colour the reference category
```{r}
df$coat_colour <- factor(df$coat_colour, levels = c("light", "mix", "dark"), ordered = FALSE)
head(df$coat_colour)
```



# 5. Generate data summary
```{r}
sumtable(df)
```



# 6. Visualise associations
## a. associtation between cortisol and sex
### i. violin plot (vioplot package) for cortisol
```{r}
par(mfrow = c(1,1))
vioplot(cortisol ~ sex, col = "firebrick",
        data = df)
```


### ii. violin plot (vioplot package) for lgCort
```{r}
par(mfrow = c(1,1))
vioplot(lgCort ~ sex, col = "lemonchiffon",
        data = df)
```


## b. association between lgCortisol and breed...
### i. violin plot (vioplot package) for lgCort
```{r}
par(mfrow = c(1,1))
vioplot(lgCort ~ breed, col = "firebrick",
        data = df)
```


### iii. stripchart for lgCort
```{r}
stripchart(lgCort ~ breed, vertical = TRUE, method = "jitter",
           col = "steelblue3", data = df, pch = 20)

```



## c. association between lgCortisol and season 
### i. violin plot (vioplot package) for lgCort
```{r}
par(mfrow = c(1,1))
vioplot(lgCort ~ season, col = "lemonchiffon",
        data = df)
```

### ii. stripchart for lgCort
```{r}
stripchart(lgCort ~ season, vertical = TRUE, method = "jitter",
           col = "steelblue3", data = df, pch = 20)

```



## d. association between lgCortisol and comorbidity 
### i. violin plot (vioplot package) for lgCort
```{r}
par(mfrow = c(1,1))
vioplot(cortisol ~ comorbidity, col = "steelblue",
        data = df)
```


### ii. violin plot (vioplot package) for lgCort
```{r}
par(mfrow = c(1,1))
vioplot(lgCort ~ comorbidity, col = "steelblue",
        data = df)
```





## e. association between cortisol vs age 
### iv. basic dot-plot for cortisol vs age
```{r}
plot(cortisol ~ age, col = "steelblue3", data = df, pch = 20)
```


### iv. basic dot-plot for lgCort vs age
```{r}
plot(lgCort ~ age, col = "steelblue3", data = df, pch = 20)
```




# STANDARDIZE DATA FOR MODELLING

## 1. standardize data for Bayesian modelling
## a. fat
```{r}
df$sFat <- standardize(df$fat_percent)
summary(df$sFat)
```

## b. age
```{r}
df$sAge <- standardize(df$age)
summary(df$sAge)
```


## c. cortisol
```{r}
df$slgCort <- standardize(df$lgC)
summary(df$slgCort)
hist(df$slgCort)
```



## 2. create dataset only containing complete data
```{r}
df2 <- na.omit(df)
```





### Combined dot plot and boxplot for LgCort vs comorbidity
```{r}
  # plot
  ggplot(data = df2, aes(x = comorbidity)) +
    geom_boxplot(aes(y = slgCort),
                 color = "black", width=0.5, lwd = 1) +
 
   geom_point(aes(y = slgCort), 
             size = 4, color = "purple",
             position = position_jitter(width = 0.2)) +
             scale_shape(solid = FALSE) +

   theme_bw() +
    labs(title = "Log hair cortisol by comorbidity") +
         labs(y = paste0("Log Hair Cortisol (standardised)")) +
         labs(x = paste0("Comorbidity")) +
         theme(axis.title.y = element_text(size=16, face = "bold"), 
               axis.title.x = element_text(size=16, face = "bold"),
               axis.text = element_text(size = 14, face = "bold"),
               title = element_text(size=18, face="bold"),
               plot.title = element_text(hjust = 0.5)) 
```








# MODEL FOR COMORBIDITY ALLOWING UNEQUEAL VARIANCE FOR COMORBIDITY

##  i. Model code

model <- brm(slgCort ~ comorbidity + sAge + breed + sex + (1 | visit),
                   family = skew_normal(),
                   data = df)

## ii. Check what priors need to be set
```{r}
default_prior(slgCort ~ comorbidity + sAge + breed + sex + (1 | visit),
                   family = skew_normal(),
                   data = df)
```





## 1. Set priors
NB skip the sigma parameter as this will be determined by the model
```{r}
# Set individual priors
prior_int <- set_prior("normal(0, 0.5)", class = "Intercept")
prior_b <- set_prior("normal(0, 1)", class = "b")
prior_b_sex <- set_prior("normal(0, 1)", class = "b", coef = "sexMale")
prior_b_co <- set_prior("normal(0.25, 1)", class = "b", coef = "comorbidityyes")
prior_b_sAge <- set_prior("normal(0, 0.5)", class = "b", coef = "sAge")
prior_sd <- set_prior("normal(0, 1)", class = "sd")
prior_alpha <- set_prior("normal(4, 2)", class = "alpha")

# Combine priors into list
priors2 <- c(prior_int, prior_b, prior_b_sex, prior_b_co, prior_b_sAge, prior_sd, prior_alpha)
```

## 2. Run model2
Increased adapt_delta >0.8 (0.999 here), as had divergent transitions
```{r}
set.seed(999)
model2 <- brm(bf(slgCort ~ comorbidity + sAge + breed + sex + (1 | visit),
                 sigma ~ comorbidity),
                   family = skew_normal(),
                   prior = priors2,
                   data = df,
                   control=list(adapt_delta = 0.99999, stepsize = 0.001, max_treedepth =15),
                   iter = 8000, warmup = 2000,
                   cores = 4,
                   save_pars = save_pars(all =TRUE),
                   sample_prior = TRUE)
```


## 3. get summary of model
```{r}
summary(model2)
```


## 4. MCMC diagnostics
## a. check distriitions and trace plots
```{r}
plot(model2)
```
Looking for hairy caterpillars


## b. try a trank plot as well
```{r}
mcmc_plot(model2, type = 'rank_overlay')
```




## 5. calculate 97% HPDI for comorbiity
Usually better than the compatability intervals given in the summary
```{r}
draws <- as.matrix(model2)
HPDI(draws[,3], 0.97) # 3rd column is draws for comorbidity
```



# 6. Calculate R2 for model
```{r}
bayes_R2(model2, probs = c(0.015, 0.5, 0.985)) # 0.015, 0.5, 0.985 are the quantiles
loo_R2(model2, probs = c(0.015, 0.5, 0.985)) # 0.015, 0.5, 0.985 are the quantiles
```




# CHECKS ON MODEL

## 1. basic check of simulations based on posterior distribution, versus the real data distribution
checks whether actual data is similar to simulated data.
```{r}
pp_check(model2, ndraws = 100) 
```



## 2. check some individual draws versus observed using pp_check
```{r}
par(mfrow = c(1,1))
pp_check(model2, type = "hist", ndraws = 11, binwidth = 0.25) # separate histograms of 11 MCMC draws vs actual data
```


## 3. other pp_check graphs
```{r}
pp_check(model2, type = "error_hist", ndraws = 11, , binwidth = 0.25) # separate histograms of errors for 11 draws
pp_check(model2, type = "scatter_avg", ndraws = 100) # scatter plot
pp_check(model2, type = "stat_2d") #  scatterplot of joint posteriors
```


## 4 error scatter average  plot (recommended by Buerkner)
```{r}
pp_check(model2, type = "error_scatter_avg")
```



## 4. pairs plot
```{r}
pairs(model2)
```



## 5. try the PSIS LOO-CV procedure to check model performance
### a. run the loo function
```{r}
loo_model2 <- loo(model2, moment_match = TRUE)
loo_model2
```



# AUTOMATED PRIOR SENSITIVITY USING THE PRIOR SENSE PACKAGE
## 1. Try automated prior sensitivity analysis using the prior sense package
First, check the sensitivity of the prior and likelihood to power-scaling. 
Posterior and posteriors resulting from power-scaling.
```{r}
powerscale_sensitivity(model2, 
                       variable = c("b_Intercept",
                                    "b_sAge",
                                    "b_breedckcs", "b_breedother", "b_breedpug", "b_breedret",
                                    "b_sexMale",
                                    "b_comorbidityyes"))
```



## 2.  Now use bayestestR package to check priors are informative
```{r}
check_prior(model2, effects = "all")
```



# CHECK PRIOR PREDICTION LINES FROM FINAL MODEL

## 1. obtain draws of priors from final model
```{r}
prior <- prior_draws(model2)
prior %>% glimpse()
```


## 2. plot prior predictions
### a. prior prediction lines for Age
```{r}
set.seed(5)

prior %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column("draw") %>% 
  expand_grid(a = c(-2, 2)) %>% 
  mutate(c = Intercept + b_sAge * a) %>% 
  
  ggplot(aes(x = a, y = c)) +
  geom_line(aes(group = draw),
            color = "firebrick", alpha = .4) +
  labs(x = "Age (std)",
       y = "log(cort) (std)") +
  coord_cartesian(ylim = c(-4, 4)) +
  theme_bw() +
  theme(panel.grid = element_blank()) 
```



### b. plot prior predictions comorbidity (yes)
#### i. line plot
```{r}
set.seed(5)

prior %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column("draw") %>% 
  expand_grid(a = c(-2, 2)) %>% 
  mutate(c = Intercept + b_comorbidityyes * a) %>% 
  
  ggplot(aes(x = a, y = c)) +
  geom_line(aes(group = draw),
            color = "firebrick", alpha = .4) +
  geom_point(color = "firebrick", size = 2) +
  labs(x = "comorbidity",
       y = "log(cort) (std)") +
  coord_cartesian(ylim = c(-4, 4)) +
  theme_bw() +
  theme(panel.grid = element_blank())
```




#### ii. combined box & whisker with dot plot for comorbidity from prior
```{r}
# set seed for repeatability
set.seed(666)

# create a new dataframe with 50 draws each from the prior
prior_50 <- prior %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column("draw") %>% 
  expand_grid(a = c(0, 1)) %>% 
  mutate(c = Intercept + b_comorbidityyes * a) %>% 
  mutate(comorbidity = ifelse(a == 0, "no", "yes"))
  

# plot
  ggplot(data = prior_50, aes(x = comorbidity)) +
    
  # add a horizontal line at y=0
    geom_hline(yintercept=0, linetype="dashed", 
                color = "grey", size=1) +
    
  # create boxplot
   geom_boxplot(aes(y = c),
                 color = "black", width=0.75, lwd = 1) +

     
  # change the y-axis limits
    ylim(-2, 2.5) +
   
  # add points
   geom_point(aes(y = c), 
             size = 4, color = "firebrick3",
             position = position_jitter(width = 0.3)) +
             scale_shape(solid = FALSE) +

   theme_bw() +
   labs(title = "Draws from prior for effect of comorbidity on hair cortisol") +
         labs(y = paste0("Log Hair Cortisol (standardised)"),
              x = paste("Comorbidity")) +  
   theme(axis.title.y = element_text(size=12, face="bold"), 

            title = element_text(size=12, face="bold"),
            plot.title = element_text(hjust = 0.5),
            axis.text.x = element_text(color = "grey50", size = 12),
            axis.text.y = element_text(color = "grey8",size = 12))
    
```


#### iii. combined box & whisker with dot plot for comorbidity from posterior
```{r}
# set seed for repeatability
set.seed(666)

draws <- as.data.frame(draws)
# create a new dataframe with 50 draws each from the prior
draws_50 <- draws %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column("draw") %>% 
  expand_grid(a = c(0, 1)) %>% 
  mutate(c = Intercept + b_comorbidityyes * a) %>% 
  mutate(comorbidity = ifelse(a == 0, "no", "yes"))
  

# plot
  ggplot(data = draws_50, aes(x = comorbidity)) +
    
 # add a horizontal line at y=0
    geom_hline(yintercept=0, linetype="dashed", 
                color = "grey", size=1) +
  
 # create boxplot
   geom_boxplot(aes(y = c),
                 color = "black", width=0.75, lwd =1) +

# change the y-axis limits
    ylim(-2, 2.5) +
  
# add points
   geom_point(aes(y = c), 
             size = 4, color = "steelblue3",
             position = position_jitter(width = 0.3)) +
             scale_shape(solid = FALSE) +
    
   theme_bw() +
   labs(title = "Draws from posterior for effect of comorbidity on hair cortisol") +
         labs(y = paste0("Log Hair Cortisol (standardised)"),
              x = paste("Comorbidity")) +  
   theme(axis.title.y = element_text(size=12, face="bold"), 

            title = element_text(size=12, face="bold"),
            plot.title = element_text(hjust = 0.5),
            axis.text.x = element_text(color = "grey50", size = 12),
            axis.text.y = element_text(color = "grey8",size = 12))

```





### c. plot prior predictions for sex (Male)
```{r}
set.seed(5)

prior %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column("draw") %>% 
  expand_grid(a = c(0, 1)) %>% 
  mutate(c = Intercept + b_sexMale * a) %>% 
  
  ggplot(aes(x = a, y = c)) +
  geom_line(aes(group = draw),
            color = "firebrick", alpha = .4) +
  geom_point(color = "firebrick", size = 2) +
  labs(x = "Sex breed",
       y = "log(cort) (std)") +
  coord_cartesian(ylim = c(-4, 4)) +
  theme_bw() +
  theme(panel.grid = element_blank()) 
```






### d. Plot prior predictions for breedckcs
```{r}
set.seed(5)

prior %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column("draw") %>% 
  expand_grid(a = c(0, 1)) %>% 
  mutate(c = Intercept + b_breedckcs * a) %>% 
  
  ggplot(aes(x = a, y = c)) +
  geom_line(aes(group = draw),
            color = "firebrick", alpha = .4) +
  geom_point(color = "firebrick", size = 2) +
  labs(x = "CKCS breed",
       y = "log(cort) (std)") +
  coord_cartesian(ylim = c(-4, 4)) +
  theme_bw() +
  theme(panel.grid = element_blank()) 
```




### e. Plot prior predictions for breedother
```{r}
set.seed(5)

prior %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column("draw") %>% 
  expand_grid(a = c(0, 1)) %>% 
  mutate(c = Intercept + b_breedother * a) %>% 
  
  ggplot(aes(x = a, y = c)) +
  geom_line(aes(group = draw),
            color = "firebrick", alpha = .4) +
  geom_point(color = "firebrick", size = 2) +
  labs(x = "Other breed",
       y = "log(cort) (std)") +
  coord_cartesian(ylim = c(-4, 4)) +
  theme_bw() +
  theme(panel.grid = element_blank()) 
```





### f. Plot prior predictions for breedpug
```{r}
set.seed(5)

prior %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column("draw") %>% 
  expand_grid(a = c(0, 1)) %>% 
  mutate(c = Intercept + b_breedpug * a) %>% 
  
  ggplot(aes(x = a, y = c)) +
  geom_line(aes(group = draw),
            color = "firebrick", alpha = .4) +
  geom_point(color = "firebrick", size = 2) +
  labs(x = "Pug breed",
       y = "log(cort) (std)") +
  coord_cartesian(ylim = c(-4, 4)) +
  theme_bw() +
  theme(panel.grid = element_blank()) 
```





### g. Plot prior predictions for breedret
```{r}
set.seed(5)

prior %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column("draw") %>% 
  expand_grid(a = c(0, 1)) %>% 
  mutate(c = Intercept + b_breedpug * a) %>% 
  
  ggplot(aes(x = a, y = c)) +
  geom_line(aes(group = draw),
            color = "firebrick", alpha = .4) +
  geom_point(color = "firebrick", size = 2) +
  labs(x = "Pug breed",
       y = "log(cort) (std)") +
  coord_cartesian(ylim = c(-4, 4)) +
  theme_bw() +
  theme(panel.grid = element_blank()) 
```







# MANUAL POSTERIOR PREDICTIVE DISTRIBUTION CHECKS
NB Uses posterior_predict

## 1. Posterior predictive distribition plots for a categorical predictor variable (comorbidity)
```{r}
# use posterior predict to simulate predictions
ppd <- posterior_predict(model2) 

par(mfrow = c(2,2))
vioplot(slgCort ~ comorbidity, data = df, main = "Observed", col = "steelblue3")
vioplot(ppd[sample(seq(1, dim(ppd)[1]), 1),] ~ df$comorbidity, main = "PPD", col = "firebrick3")
vioplot(ppd[sample(seq(1, dim(ppd)[1]), 1),] ~ df$comorbidity, main = "PPD", col = "firebrick3")
vioplot(ppd[sample(seq(1, dim(ppd)[1]), 1),] ~ df$comorbidity, main = "PPD", col = "firebrick3")
```


## 2. Posterior predictive distribition plots for a categorical predictor variable (sex)
```{r}
par(mfrow = c(2,2))
vioplot(slgCort ~ sex, data = df, main = "Observed", col = "steelblue")
vioplot(ppd[sample(seq(1, dim(ppd)[1]), 1),] ~ df$sex, main = "PPD", col = "firebrick3")
vioplot(ppd[sample(seq(1, dim(ppd)[1]), 1),] ~ df$sex, main = "PPD", col = "firebrick3")
vioplot(ppd[sample(seq(1, dim(ppd)[1]), 1),] ~ df$sex, main = "PPD", col = "firebrick3")
```


## 3. Posterior predictive distribition plots for a categorical predictor variable with small group size
```{r}
par(mfrow = c(2,2))
stripchart(slgCort ~ breed, vertical = TRUE, method = "jitter",
           col = "steelblue3", data = df, pch = 20, main = "Observed")
stripchart(ppd[sample(seq(1, dim(ppd)[1]), 1),] ~ breed, vertical = TRUE, method = "jitter",
           col = "firebrick3", data = df, pch = 20, main = "PPD")
stripchart(ppd[sample(seq(1, dim(ppd)[1]), 1),] ~ breed, vertical = TRUE, method = "jitter",
           col = "firebrick3", data = df, pch = 20, main = "PPD")
stripchart(ppd[sample(seq(1, dim(ppd)[1]), 1),] ~ breed, vertical = TRUE, method = "jitter",
           col = "firebrick3", data = df, pch = 20, main = "PPD")
```




# ANALYSING THE POSTERIOR DISTRIBUTION

## 1. plot conditional effects from model
### a. all conditional effects
```{r}
plot(conditional_effects(model2), ask = FALSE)
```

### a. conditional effect of comorbidity only
```{r}
ce <- conditional_effects(model2, effects = "comorbidity", method = "posterior_predict")
ce_df <- ce[[1]][c(1, 9:12)]

ggplot(ce_df, aes(x=comorbidity, y=estimate__, group=1)) +
    geom_errorbar(width=.1, aes(ymin=lower__, ymax=upper__), colour=c("#F8766D", "#00BFC4"), linewidth = 2) +
    geom_point(shape=21, size=10, fill=c("#F8766D", "#00BFC4")) +
   theme_bw() +
    labs(title = "Conditional effect of comorbidity on hair cortisol") +
         labs(y = paste0("Log Hair Cortisol (standardised)")) +
         labs(x = paste0("Comorbidity")) +
         theme(axis.title.y = element_text(size=12, face="bold"), 
               axis.title.x = element_text(size=12, face="bold"),
               title = element_text(size=12, face="bold"),
               plot.title = element_text(hjust = 0.5),
               axis.text.x = element_text(color = "grey50", size = 12),
               axis.text.y = element_text(color = "grey50", size = 12))
```





## 2. mcmc_plot of model
### a.just parameters of beta variables and also sigma for comorbidity
```{r}
mcmc_plot(model2,
          variable = c("b_comorbidityyes",
         "b_sAge",
         "b_breedckcs",
         "b_breedpug",
         "b_breedret",
         "b_breedother",
         "b_sigma_comorbidityyes"))
```


### b. all parameters except nu
```{r}
mcmc_plot(model2,
          variable = c("Intercept",
         "b_Intercept", "alpha",
         "b_sigma_Intercept",
         "b_comorbidityyes",
         "b_sAge",
         "b_sexMale",
         "b_breedckcs",
         "b_breedpug",
         "b_breedret",
         "b_breedother",
         "b_sigma_comorbidityyes"))
```

```{r}
model2$fit
```




### c. Parameters versus their priors
#### i. Intercept
```{r}
mcmc_plot(model2,
          variable = c("Intercept", "prior_Intercept",
         "b_Intercept"))
```



#### ii. sigma_Intercept
```{r}
mcmc_plot(model2,
          variable = c("b_sigma_Intercept", "prior_Intercept_sigma"))
```



#### iii. all betas versus their priors
```{r}
mcmc_plot(model2,
          variable = c("b_comorbidityyes", "prior_b_comorbidityyes",
                       "b_sAge", "prior_b_sAge",
                       "b_sexMale", "prior_b_sexMale",
                       "b_breedckcs", "prior_b_breedckcs",
                       "b_breedpug", "prior_b_breedpug",
                       "b_breedret", "prior_b_breedret",
                       "b_breedother", "prior_b_breedother"))
```


##### comorbidity vs prior
###### 1. distributional
```{r}
mcmc_plot(model2,
          variable = c("b_comorbidityyes", "prior_b_comorbidityyes")) +


   theme_classic() +
    labs(title = "Prior vs posterior distribution for comorbidity") +
         labs(y = "") +
         labs(x = paste0("Possible parameter values")) +
    scale_y_discrete(labels=c("prior_b_comorbidityyes" = "Prior", "b_comorbidityyes" = "Posterior"),
                     limits = c("prior_b_comorbidityyes", "b_comorbidityyes")) +
         theme(axis.title.y = element_text(size=12, face="bold"), 
               axis.title.x = element_text(size=12, face="bold"),
               title = element_text(size=12, face="bold"),
               plot.title = element_text(hjust = 0.5),
               axis.text.x = element_text(color = "grey50", size = 12),
               axis.text.y = element_text(color = "grey8",size = 12))

```


###### 2. density
```{r}
mcmc_plot(model2,
          variable = c("b_comorbidityyes", "prior_b_comorbidityyes"),
          type = "areas") +

   theme_classic() +
    labs(title = "Prior vs posterior distribution for comorbidity") +
         labs(y = "") +
         labs(x = paste0("Possible parameter values")) +
    scale_y_discrete(labels=c("prior_b_comorbidityyes" = "Prior", "b_comorbidityyes" = "Posterior"),
                     limits = c("prior_b_comorbidityyes", "b_comorbidityyes")) +
         theme(axis.title.y = element_text(size=12, face="bold"), 
               axis.title.x = element_text(size=12, face="bold"),
               title = element_text(size=12, face="bold"),
               plot.title = element_text(hjust = 0.5),
               axis.text.x = element_text(color = "grey50", size = 12),
               axis.text.y = element_text(color = "grey8",size = 12))
```




#### iv. sAge
```{r}
mcmc_plot(model2,
          variable = c("b_sAge", "prior_b_sAge"))
```


#### v. sexMale
```{r}
mcmc_plot(model2,
          variable = c("b_sexMale", "prior_b_sexMale"))
```

#### vi. breedckcs
```{r}
mcmc_plot(model2,
          variable = c("b_breedckcs", "prior_b_breedckcs"))
```


#### v. breedpug
```{r}
mcmc_plot(model2,
          variable = c("b_breedpug", "prior_b_breedpug"))
```



#### vi. breedret
```{r}
mcmc_plot(model2,
          variable = c("b_breedret", "prior_b_breedret"))
```



#### vii. breedother
```{r}
mcmc_plot(model2,
          variable = c("b_breedother", "prior_b_breedother"))
```










## 3. plot all posterior distributions
```{r}
posterior <- as.matrix(model2)
mcmc_areas(posterior,
    pars = c("b_Intercept",
         "b_sigma_Intercept",
         "alpha",
         "b_comorbidityyes",
         "b_sAge",
         "b_sexMale",
         "b_breedckcs",
         "b_breedpug",
         "b_breedret",
         "b_breedother",
         "b_sigma_comorbidityyes"),
    prob = 0.75) # arbitrary threshold for shading probability mass
```




## 4. plot posterior distribution for comorbidity
```{r}
posterior <- as.matrix(model2)
mcmc_areas(posterior,
    pars = c("b_comorbidityyes",
             "b_sigma_comorbidityyes"),
    prob = 0.97) +  # arbitrary threshold for shading probability mass 
 
   theme_classic() +
     labs(title = "Posterior distribution for comorbidity parameters", 
         y = "Density distribution", 
         x = "Possible parameter values") +
    scale_y_discrete(labels=c("b_comorbidityyes" = "Beta", "b_sigma_comorbidityyes" = "Sigma"),
                     limits=c("b_sigma_comorbidityyes","b_comorbidityyes")) +
         theme(axis.title.y = element_text(size=12, face="bold"), 
               axis.title.x = element_text(size=12, face="bold"),
               title = element_text(size=12, face="bold"),
               plot.title = element_text(hjust = 0.5),
               axis.text.x = element_text(color = "grey50", size = 12),
               axis.text.y = element_text(color = "grey8",size = 12))

```


 
## 5. Describe the posterior visually
### a. all beta parameters
```{r}
# Focus on describing posterior
hdi_range <- hdi(model2, ci = c(0.65, 0.70, 0.80, 0.89, 0.95))
plot(hdi_range, show_intercept = T)
```



### b. beta for comorbidity only

```{r}
# Focus on describing posterior
hdi_range <- hdi(model2, ci = c(0.65, 0.70, 0.80, 0.89, 0.97), parameters = "comorbidityyes")
plot(hdi_range, show_intercept = T) +

    labs(title = "Posterior distribution for comorbidity") +
         labs(y = "Density distribution") +
         labs(x = "Possible parameter values") +
           theme(axis.title.y = element_text(size=12, face="bold"), 
               axis.title.x = element_text(size=12, face="bold"),
               title = element_text(size=12, face="bold"),
               plot.title = element_text(hjust = 0.5),
               axis.text.x = element_text(color = "grey50", size = 12),
               axis.text.y = element_text(color = "grey8",size = 12))
```




# POSTERIOR PREDICTION

## 1. posterior prediction for effect of comorbidity on Log cortisol
(when visit = vO, breed = mix, sAge = 0 (mean), sex = Female)
```{r}
# create new dataframe which contains results of the 10th dog
new_data <- rbind(df2[10,], df2[10,])
# Now change one category to be different
new_data$comorbidity <- c("yes", "no")
# Visualise df to make sure it has worked
new_data


# Now get mean predictions from the draws of the model
pred_means <- posterior_predict(model2, newdata = new_data)


# Compare difference in means for each breed versus mix
difference <- pred_means[,1] - pred_means[,2]

par(mfrow = c(2,2))

# Examine mean of difference for comorbidity
mean(difference)
# View histogram of this
dens(difference)
# Create HPDI
HPDI(difference, 0.97)

```




# HYPOTHESIS TESTS

## 1. hypothesis test to check if mean association between cortisol and comborbidity is >0
```{r}
draws <- as.matrix(model2)
mean(draws[,3] >0)
```


## 2. check 97% credible interval of with HPDI for comorbidity from draws 
```{r}
HPDI(draws[,3], prob=0.97)
```



## 3. hypothesis test that the effect of comorbidity is positive
```{r}
hypothesis(model2, "comorbidityyes >0")
hypothesis(model2, "comorbidityyes <0")
```



# VISUALISING PREDICTIONS
## 1. visualising the posterior predictions from the model
```{r}
# create new dataframe which contains results of all dogs
new_data1 <- df2
# Now change one category to be different
new_data1$comorbidity <- c("yes")

# create a second new dataframe which contains results of all dogs
new_data2 <- df2
# Now change one category to be different
new_data2$comorbidity <- c("no")

# Now get predictions from the draws of the models
pred_nd1 <- posterior_predict(model2, newdata = new_data1)
pred_nd2 <- posterior_predict(model2, newdata = new_data2)
pred_diff <- pred_nd1 - pred_nd2
pred_diff <- data.frame(pred_diff)

# Create mean of differences for each column (dog) of the dataframe
pred_diff_mean <- apply(pred_diff, 2, mean)
# View histogram of mean differences
hist(pred_diff_mean)
# mean difference
mean(pred_diff_mean)
# Create HPDI
HPDI(pred_diff_mean, 0.97)
```


## 2. make predictions of log cortisol for each dog and compare with actual data
```{r}
pred_slgCort <- posterior_epred(model2)
av_pred_slgCort <- colMeans(pred_slgCort)
plot(av_pred_slgCort ~ df$slgCort)
```






## 3. plot the counterfactual effect of "do comorbidity" on slgCort
### a. plot estimates and 95% credible intervals 
```{r}
set.seed(666)
nd <- tibble(comorbidity = c('no', 'yes'), visit = "v0", sAge = 0, sex = "Female", breed = "mix", )

p1 <-
  predict(model2,
          resp = "slgCort",
          newdata = nd) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  
  ggplot(aes(x = comorbidity, y = Estimate, ymin = Q2.5, ymax = Q97.5)) +
  
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5),
                 linewidth = 2, color = c("#F8766D", "#00BFC4"), alpha = 3/5) +
  geom_point(size = 8, color = c("#F8766D", "#00BFC4")) +

   theme_bw() +
    labs(title = "Log hair cortisol (standardised)") +
         labs(y = paste0("Log hair cortisol (standardised)")) +
         labs(x = paste0("Comorbidity")) +
         theme(axis.title.y = element_text(size=12, face="bold"), 
               axis.title.x = element_text(size=12, face="bold"),
               title = element_text(size=12, face="bold"),
               plot.title = element_text(hjust = 0.5))

plot(p1)
```



# Citations and package versions for packages
```{r}
citation()

citation("dlookr")
packageVersion("dlookr")

citation("dplyr")
packageVersion("dplyr")

citation("ggplot2")
packageVersion("ggplot2")

citation("psych")
packageVersion("psych")

citation("reshape")
packageVersion("reshape")

citation("readxl")
packageVersion("readxl")

citation("tidyverse")
packageVersion("tidyverse")

citation("vioplot")
packageVersion("vioplot")

citation("vtable")
packageVersion("vtable")


citation("brms")
packageVersion("brms")

citation("bayesplot")
packageVersion("bayesplot")

citation("bayestestR")
packageVersion("bayestestR")

citation("rethinking")
packageVersion("rethinking")

citation("loo")
packageVersion("loo")

citation("priorsense")
packageVersion("priorsense")

```



